# Trial 1 : Integrated analysis (HD+patient)

## Processing

```{r}
library(Seurat)
BiocManager::install("dittoSeq")
BiocManager::install("harmony")
library(dittoSeq)
library(harmony)
source('/home/sbm/kyungmin/Rstudio/RunHarmony_2.R')
```

```{r}
set.seed(123)
setwd('/home/sbm/kyungmin/HYIBB/scRNA/Data/GSE210065/Result/')

CHOL_seurat_cut <- readRDS('CHOL_seurat_cut.rds') # contains protein&RNA data of HD+patient
```

```{r}
# Retrieve only healthy and patient before treatment for initial clustering
CHOL_HD_patient_after <- subset(CHOL_seurat_cut, subset = timepoint %in% c("HD", "T02", "T03"))
```

## Initial clustering based on Protein

### Processing

```{r}
# Normalize ADT data (CLR method recommended)
DefaultAssay(CHOL_HD_patient_after) <- "ADT"
CHOL_HD_patient_after <- NormalizeData(CHOL_HD_patient_after, normalization.method = "CLR", margin = 2, assay = "ADT")
CHOL_HD_patient_after <- ScaleData(CHOL_HD_patient_after, assay = "ADT")
```

### Clustering

```{r}
# Cluster directly on protein levels

DefaultAssay(CHOL_HD_patient_after) <- "ADT"
CHOL_HD_patient_after <- RunPCA(CHOL_HD_patient_after, features = rownames(CHOL_HD_patient_after), reduction.name = "pca_adt", reduction.key = "pca_adt_", verbose = FALSE)

DimPlot(CHOL_HD_patient_after, reduction = "pca_adt", group.by= "patient")
ElbowPlot(CHOL_HD_patient_after , ndims = 50)

CHOL_HD_patient_after <- RunUMAP(CHOL_HD_patient_after, reduction = "pca_adt", reduction.key = "adt_umap_", reduction.name = "umap_adt" , dims = 1:15)
CHOL_HD_patient_after <- FindNeighbors(CHOL_HD_patient_after, features = rownames(CHOL_HD_patient_after), dims = NULL)
CHOL_HD_patient_after <- FindClusters(CHOL_HD_patient_after, resolution = 0.1, graph.name = "ADT_snn")

# DimPlot(CHOL_HD_patient_B, reduction = "umap_adt")

# CHOL_seurat_cut[["adt_snn"]] <- FindNeighbors(CHOL_seurat_cut, dims = 1:20)
# CHOL_seurat_cut <- FindClusters(CHOL_seurat_cut, resolution = 0.4, graph.name = "adt_snn")

DimPlot(CHOL_HD_patient_after, reduction = "umap_adt", pt.size = 0.05, label = T)
# tmp <- subset(HD_Tcell_Naive_C0, downsample = 300)
protein_markers_HD_patient_after <- FindAllMarkers(CHOL_HD_patient_after, assay = "ADT", , only.pos = TRUE)


# heatmap <- DoHeatmap(CHOL_seurat_cut, features = unique(protein_markers$gene), slot="scale.data", assay="ADT", size=3) + theme(text=element_text(size=10))

heatmap <- DoHeatmap(subset(CHOL_HD_patient_after, downsample = 1000), features = unique(protein_markers_HD_patient_after$gene), assay = "ADT", angle = 90) + NoLegend()

RNA_markers_HD_patient_after <- FindAllMarkers(CHOL_HD_patient_after, assay = "RNA", logfc.threshold = 0.10, only.pos = TRUE, min.diff.pct = 0.2)

CHOL_HD_patient_after[["ADT_res0.1.pc15"]] <- Idents(CHOL_HD_patient_after)
```

### Visualization : Check plots

```{r}
# check gene feature plot/violin plot

protein_gene_name = c("CD95-FAS", "CD11a-ITGAL", "CD183-CXCR3", "CD127-IL7R", "CD62L-SELL", "CD28-CD28", "CD197-CCR7", "CD27-CD27", "CD5-CD5", "CD49d-ITGA4", "CD279-PDCD1", "CD4-CD4", "TCRab-TRA-TRB", "CD45RA-PTPRC", "CD8-CD8A", "CD3-CD3E", "CD45RO-PTPRC")



Idents(CHOL_HD_patient_after) <- "ADT_res0.2.pc15"

DefaultAssay(CHOL_HD_patient_after) <- "ADT"
DimPlot(CHOL_HD_patient_after, reduction = "umap_adt", pt.size = 0.05, label = T)
FeaturePlot(CHOL_HD_patient_after, protein_gene_name, slot = "data" , ncol =5)
VlnPlot(CHOL_HD_patient_after, protein_gene_name, group.by = "ADT_res0.2.pc15", pt.size =0, slot = "data" , ncol = 5)
```

## Run Harmony

```{r}
CHOL_HD_patient_after <- RunHarmony.Seurat(CHOL_HD_patient_after, "patient", 
                                               reduction.use  = "pca_adt",
                                               reduction.save = "harmony_adt",
                                               dims.use = 1:30)
```

```{r}
CHOL_HD_patient_after <- RunUMAP(CHOL_HD_patient_after, reduction = "harmony_adt", reduction.key = "adt_harmony_umap_", reduction.name = "harmony_umap_adt" , dims = 1:30)


CHOL_HD_patient_after <- FindNeighbors(CHOL_HD_patient_after, 
                                           features = rownames(CHOL_HD_patient_after)
                                           , dims = 1:30,
                                           reduction = "harmony_adt",
                                           graph.name = "ADT_harmony_snn")

CHOL_HD_patient_after <- FindClusters(CHOL_HD_patient_after, 
                                          resolution = 0.25,
                                          graph.name = "ADT_harmony_snn")
```

### Visualization after harmony

```{r}
DimPlot(CHOL_HD_patient_after, reduction = "harmony_umap_adt")
DimPlot(CHOL_HD_patient_after, reduction = "harmony_umap_adt", group.by = "timepoint")
DimPlot(CHOL_HD_patient_after, reduction = "harmony_umap_adt", split.by = "timepoint")
FeaturePlot(CHOL_HD_patient_after, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo", "S.Score","G2M.Score") , coord.fixed = T, ncol= 3, reduction = "harmony_umap_adt")
DimPlot(CHOL_HD_patient_after, reduction = "harmony_umap_adt", split.by = "patient", ncol = 6)
```

```{r}
pt <- table(CHOL_HD_patient_after$seurat_clusters, CHOL_HD_patient_after$patient)
pt <- as.data.frame(pt)
pt$Var1 <- as.character(pt$Var1)

# pt$Var2 <- factor(pt$Var2, levels = c("HD" , "T02", "T03"))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  #scale_fill_manual(values = brewer.pal(12, "Paired")) +
  scale_fill_manual(values = hue_pal()(9)) +
  theme(legend.title = element_blank())
```

```{r}
protein_gene_name = c("CD95-FAS", "CD11a-ITGAL", "CD183-CXCR3", "CD127-IL7R", "CD62L-SELL", "CD28-CD28", "CD197-CCR7", "CD27-CD27", "CD5-CD5", "CD49d-ITGA4", "CD279-PDCD1", "CD4-CD4", "TCRab-TRA-TRB", "CD45RA-PTPRC", "CD8-CD8A", "CD3-CD3E", "CD45RO-PTPRC" , "CD123-IL3RA", "CD235a-b-GYPA-GYPB")

# heatmap after harmony
protein_markers_HD_patient_after <- FindAllMarkers(CHOL_HD_patient_after, assay = "ADT", , only.pos = TRUE)

DoHeatmap(subset(CHOL_HD_patient_after, downsample = 1000), features = unique(protein_markers_HD_patient_after$gene), assay = "ADT", angle = 90) + NoLegend()

cluster_marker_name <- protein_markers_HD_patient_after$gene

Idents(CHOL_HD_patient_after) <- "ADT_harmony_snn_res.0.25"

DefaultAssay(CHOL_HD_patient_after) <- "ADT"
DimPlot(CHOL_HD_patient_after, reduction = "harmony_umap_adt", pt.size = 0.05, label = T)
FeaturePlot(CHOL_HD_patient_after, protein_gene_name, slot = "data" ,reduction = "harmony_umap_adt", ncol =5)
VlnPlot(CHOL_HD_patient_after, protein_gene_name, group.by = "ADT_harmony_snn_res.0.25", pt.size =0, slot = "data" , ncol = 5)

# dittoDimPlot(CHOL_HD_patient_after, "DICE", reduction.use = "harmony_umap_adt", do.letter = FALSE, do.label=F, split.by = "DICE")
```

# Re-clustering (subset 0,3 clusters)

```{r}
Tcell_48_after <- subset(CHOL_HD_patient_after, subset = (ADT_harmony_snn_res.0.25 ==0 |  ADT_harmony_snn_res.0.25==3) )
```

## Initial clustering based on **RNA**

```{r}
DefaultAssay(Tcell_48_after) <- "RNA"
```

### Variable Feature Selection (consider batch)

```{r}
# normalize RNA count data first
Tcell_48_after <- NormalizeData(Tcell_48_after)

# find variable features per patient
Tcell_48_after.list <- SplitObject(Tcell_48_after, split.by = "patient")
for (i in 1:length(Tcell_48_after.list)) {
  Tcell_48_after.list[[i]] <- FindVariableFeatures(Tcell_48_after.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = TRUE)
}

# Return top scoring 2,000 features
features_Tcell_48_after <- SelectIntegrationFeatures(Tcell_48_after.list)

rm(Tcell_48_after.list)
```

### Scale data

```{r}
Tcell_48_after <- ScaleData(Tcell_48_after, features = features_Tcell_48_after, vars.to.regress = c("nCount_RNA", "S.Score", "G2M.Score"))
```

### Run PCA

```{r}
Tcell_48_after = RunPCA(Tcell_48_after, npcs = 40, verbose = TRUE , features = features_Tcell_48_after)

ElbowPlot(Tcell_48_after, ndims = 40, reduction = "pca")
VizDimLoadings(Tcell_48_after, dims = 1:4, reduction = "pca", ncol = 2)

DimPlot(Tcell_48_after, reduction = "pca", na.value = "grey50" , group.by = "patient")
DimPlot(Tcell_48_after, reduction = "pca", na.value = "grey50" , group.by = "batch")
```

### Run Umap

```{r}
Tcell_48_after <- FindNeighbors(Tcell_48_after, dims = 1:30)
Tcell_48_after <- FindClusters(Tcell_48_after, resolution = 0.4)
Tcell_48_after <- RunUMAP(Tcell_48_after, dims = 1:30)
```

### Visualization : Check plots

```{r}
# check plot

DimPlot(Tcell_48_after, reduction = "umap", pt.size = 0.05)

DimPlot(Tcell_48_after, reduction = "umap", pt.size = 0.05 , group.by = "patient")
DimPlot(Tcell_48_after, reduction = "umap", pt.size = 0.05 , group.by = "batch")

# dittoDimPlot(C0_healthy, "DICE", reduction.use = "umap", do.letter = FALSE, do.label=F, split.by = "DICE")

FeaturePlot(Tcell_48_after, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo", "S.Score","G2M.Score") , coord.fixed = T, ncol= 3)
```

### Run Harmony

```{r}
Tcell_48_after <- RunHarmony.Seurat(Tcell_48_after, "patient", 
                                          reduction.use  = "pca",
                                          reduction.save = "harmony",
                                          dims.use = 1:30)


Tcell_48_after <- Tcell_48_after %>%
    RunUMAP(reduction = "harmony", dims= 1:30)



Tcell_48_after <- Tcell_48_after %>%
  FindNeighbors(reduction = "harmony" , dims = 1:30) %>%
    FindClusters(resolution = 0.2) 
```

### Visualization after harmony

```{r}
DimPlot(Tcell_48_after, reduction = "umap", pt.size = 0.05)

DimPlot(Tcell_48_after, reduction = "umap", pt.size = 0.05 , group.by = "patient")
DimPlot(Tcell_48_after, reduction = "umap", pt.size = 0.05 , group.by = "timepoint")

# dittoDimPlot(C0_healthy, "DICE", reduction.use = "umap", do.letter = FALSE, do.label=F, split.by = "DICE")

FeaturePlot(Tcell_48_after, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo", "S.Score","G2M.Score") , coord.fixed = T, ncol= 3, reduction = "umap")


FeaturePlot(Tcell_48_after, features = c("ADGRE5", "TCF7", "BCL2", "IL7R") , coord.fixed = T, ncol= 3, reduction = "umap")

VlnPlot(Tcell_48_after, c("ADGRE5", "TCF7", "BCL2", "IL7R"), group.by =  "timepoint", pt.size=0)

VlnPlot(Tcell_48_after, c("ADGRE5", "TCF7", "BCL2", "IL7R"), group.by =  "seurat_clusters", pt.size=0)

FeaturePlot(Tcell_48_after, features = c("IFNG", "IL13", "IL4", "IL5", "IL17A", "CXCR5", "PDCD1") , coord.fixed = T, ncol= 4, reduction = "umap")

VlnPlot(Tcell_48_after, c("IFNG", "IL13", "IL4", "IL5", "IL17A", "CXCR5", "PDCD1"), group.by =  "seurat_clusters", pt.size=0, ncol=4)
```

### Marker selection

```{r}
# marker selection from naive from cancer
Tcell_48_after_RNA_marker <- FindAllMarkers(Tcell_48_after, 
                                           logfc.threshold = 0.1,
                                           only.pos = TRUE,
                                           min.pct = 0.1,
                                           assay = "RNA")
```

```{r}
Tcell_48_after_protein_marker <- FindAllMarkers(Tcell_48_after, 
                                           logfc.threshold = 0.1,
                                           only.pos = TRUE,
                                           min.pct = 0.1,
                                           assay = "ADT")
```

```{r}
setwd('/home/sbm/kyungmin/HYIBB/scRNA/Data/GSE210065/Result3(after_treatment)/')
write.csv(Tcell_48_after_RNA_marker, "Tcell_48_after_RNA_marker.csv")
write.csv(Tcell_48_after_protein_marker, "Tcell_48_after_protein_marker.csv")
```

### Heatmap

```{r}
# heatmap
Tcell_48_after_RNA_marker %>%  group_by(cluster) %>%  top_n(n = 10, wt = avg_log2FC) -> top10_RNA_Tcell_markers

# gene hard clustering
top_Tcell_markers <- Tcell_48_after_RNA_marker %>%
  group_by(gene) %>%
  top_n(n = 1, wt = avg_log2FC) %>%
  ungroup()

top_Tcell_markers %>%  group_by(cluster) %>%  top_n(n = 10, wt = avg_log2FC) -> top10_RNA_hard_Tcell_markers


heatmap1_c <- DoHeatmap(subset(Tcell_48_after, downsample = 4000), features = top10_RNA_Tcell_markers$gene, slot="scale.data", assay="RNA", size=3) + theme(text=element_text(size=10))

# hard-clustering
heatmap2_c <- DoHeatmap(subset(Tcell_48_after, downsample = 4000), features = top10_RNA_hard_Tcell_markers$gene, slot="scale.data", assay="RNA", size=3) + theme(text=element_text(size=10))
```

### Check cluster proportion

```{r}
pt <- table(Tcell_48_after$seurat_clusters, Tcell_48_after$response)
pt <- as.data.frame(pt)
pt$Var1 <- as.character(pt$Var1)

# pt$Var2 <- factor(pt$Var2, levels = c("T01", "HD"))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  #scale_fill_manual(values = brewer.pal(12, "Paired")) +
  scale_fill_manual(values = hue_pal()(length(unique(Tcell_after$seurat_clusters))) )   +
  theme(legend.title = element_blank())
```

### Create excel sheet (cell number)

```{r}
library(Seurat)
library(openxlsx)

# create excel file
wb <- createWorkbook()

# extract metadata
metadata_Tcell_48_after <- Tcell_48_after@meta.data

# extract unique timepoint
timepoints <- unique(metadata_Tcell_48_after$timepoint)

# repeat for each timepoints
for (tp in timepoints) {
  # filter specific timepoint
  data_tp <- metadata_Tcell_48_after[metadata_Tcell_48_after$timepoint == tp, ]
  
  # create (number of cells per cluster/patients)
  freq_table <- table(data_tp$patient, data_tp$seurat_clusters)
  
  # transform into dataframe
  df <- as.data.frame.matrix(freq_table)
  
  # add data to new sheet
  addWorksheet(wb, sheetName = paste0("Timepoint_", tp))
  writeData(wb, sheet = paste0("Timepoint_", tp), df)
}

# save excel file
saveWorkbook(wb, "/home/sbm/kyungmin/HYIBB/scRNA/Data/GSE210065/Result3(after_treatment)/Tcell_48_after_cluster_distribution.xlsx", overwrite = TRUE)
```

# Re-clustering (cluster 1, 2)

```{r}
# subset C1 & C2
naive_CD4_after <- subset(CHOL_HD_patient_after, subset = (ADT_harmony_snn_res.0.25 ==1 |  ADT_harmony_snn_res.0.25==2) )
```

## Clustering

### HVG selection

```{r}
DefaultAssay(naive_CD4_after) <- "RNA"
naive_CD4_after <- NormalizeData(naive_CD4_after)

# find variable features per batch
naive_CD4_after.list <- SplitObject(naive_CD4_after, split.by = "patient")
for (i in 1:length(naive_CD4_after.list)) {
  naive_CD4_after.list[[i]] <- FindVariableFeatures(naive_CD4_after.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = TRUE)
}

# Return top scoring 2,000 features
features_naive_after <- SelectIntegrationFeatures(naive_CD4_after.list)

rm(naive_CD4_after.list)
```

### Scale data

```{r}
naive_CD4_after <- ScaleData(naive_CD4_after, 
                           features = features_naive_after, 
                           vars.to.regress= c("nCount_RNA", "S.Score", "G2M.Score")
                           )
```

### Run PCA

```{r}
naive_CD4_after = RunPCA(naive_CD4_after, npcs = 40, verbose = TRUE , features = features_naive_after)

ElbowPlot(naive_CD4_after, ndims = 40, reduction = "pca")
VizDimLoadings(naive_CD4_after, dims = 1:4, reduction = "pca", ncol = 2)

DimPlot(naive_CD4_after, reduction = "pca", na.value = "grey50" , group.by = "patient")
DimPlot(naive_CD4_after, reduction = "pca", na.value = "grey50" , group.by = "batch")
```

### Run Umap

```{r}
naive_CD4_after <- FindNeighbors(naive_CD4_after, dims = 1:30)
naive_CD4_after <- FindClusters(naive_CD4_after, resolution = 0.4)
naive_CD4_after <- RunUMAP(naive_CD4_after, dims = 1:30)
```

### Visualization

```{r}
DimPlot(naive_CD4_after, reduction = "umap", pt.size = 0.05)

DimPlot(naive_CD4_after, reduction = "umap", pt.size = 0.05 , group.by = "patient")
DimPlot(naive_CD4_after, reduction = "umap", pt.size = 0.05 , group.by = "batch")

# dittoDimPlot(naive_CD4_after, "DICE", reduction.use = "umap", do.letter = FALSE, do.label=F, split.by = "DICE")

FeaturePlot(naive_CD4_after, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo", "S.Score","G2M.Score") , coord.fixed = T, ncol= 3)
```

## Batch correction for naive CD4+ T from cancer patients

## Run Harmony

```{r}
naive_CD4_after <- RunHarmony.Seurat(naive_CD4_after, "patient", 
                                          reduction.use  = "pca",
                                          reduction.save = "harmony",
                                          dims.use = 1:30)


naive_CD4_after <- naive_CD4_after %>%
    RunUMAP(reduction = "harmony", dims= 1:30, n.neighbors = 20, min.dist = 0.2)



naive_CD4_after <- naive_CD4_after %>%
  FindNeighbors(reduction = "harmony" , dims = 1:30) %>%
    FindClusters(resolution = 0.2) 
```

```{r}
# naive_CD4_umap_test <- FindNeighbors(naive_CD4, dims = 1:30)
# naive_CD4_umap_test <- FindClusters(naive_CD4_umap_test, resolution = 0.4)
# naive_CD4_umap_test <- RunUMAP(naive_CD4_umap_test, dims = 1:30, 
#                                n.neighbors = 20, min.dist = 0.2)
```

### Visualization after harmony

```{r}
DimPlot(naive_CD4_after, reduction = "umap", pt.size = 0.05)

DimPlot(naive_CD4_after, reduction = "umap", pt.size = 0.05 , group.by = "patient")
DimPlot(naive_CD4_after, reduction = "umap", pt.size = 0.05 , group.by = "timepoint")

# dittoDimPlot(C0_healthy, "DICE", reduction.use = "umap", do.letter = FALSE, do.label=F, split.by = "DICE")

FeaturePlot(naive_CD4_after, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo", "S.Score","G2M.Score") , coord.fixed = T, ncol= 3, reduction = "umap")


FeaturePlot(naive_CD4_after, features = c("ADGRE5", "TCF7", "BCL2", "IL7R") , coord.fixed = T, ncol= 3, reduction = "umap")

VlnPlot(naive_CD4_after, c("ADGRE5", "TCF7", "BCL2", "IL7R"), group.by =  "timepoint", pt.size=0)

VlnPlot(naive_CD4_after, c("ADGRE5", "TCF7", "BCL2", "IL7R"), group.by =  "seurat_clusters", pt.size=0)
```

### Check Proportion

```{r}
pt <- table(naive_CD4_after$seurat_clusters, naive_CD4_after$response)
pt <- as.data.frame(pt)
pt$Var1 <- as.character(pt$Var1)

# pt$Var2 <- factor(pt$Var2, levels = c("T01", "HD"))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  #scale_fill_manual(values = brewer.pal(12, "Paired")) +
  scale_fill_manual(values = hue_pal()(length(unique(naive_CD4_after$seurat_clusters))) )   +
  theme(legend.title = element_blank())
```

```{r}
# Retrieve proportion

naive_CD4_after_proportion <- as.data.frame(table(naive_CD4_after$patient, naive_CD4_after$seurat_clusters,  naive_CD4_after$timepoint))
```

### Create excel sheet (cell number)

```{r}
library(Seurat)
library(openxlsx)

# create excel file
wb <- createWorkbook()

# extract metadata
metadata_naive_CD4 <- naive_CD4_after@meta.data

# extract unique timepoint
timepoints <- unique(metadata_naive_CD4$timepoint)

# repeat for each timepoints
for (tp in timepoints) {
  # filter specific timepoint
  data_tp <- metadata_naive_CD4[metadata_naive_CD4$timepoint == tp, ]
  
  # create (number of cells per cluster/patients)
  freq_table <- table(data_tp$patient, data_tp$seurat_clusters)
  
  # transform into dataframe
  df <- as.data.frame.matrix(freq_table)
  
  # add data to new sheet
  addWorksheet(wb, sheetName = paste0("Timepoint_", tp))
  writeData(wb, sheet = paste0("Timepoint_", tp), df)
}

# save excel file
saveWorkbook(wb, "/home/sbm/kyungmin/HYIBB/scRNA/Data/GSE210065/Result3(after_treatment)/naive_CD4_after_cluster_distribution.xlsx", overwrite = TRUE)
```

### Marker selection

```{r}
# marker selection from naive from cancer
naive_CD4_after_RNA_marker <- FindAllMarkers(naive_CD4_after, 
                                           logfc.threshold = 0.1,
                                           only.pos = TRUE,
                                           min.pct = 0.1,
                                           assay = "RNA")
```

### Heatmap

```{r}
# heatmap
naive_CD4_after_RNA_marker %>%  group_by(cluster) %>%  top_n(n = 10, wt = avg_log2FC) -> top10_RNA_naive_CD4_markers

# gene hard clustering
top_naive_CD4_markers <- naive_CD4_after_RNA_marker %>%
  group_by(gene) %>%
  top_n(n = 1, wt = avg_log2FC) %>%
  ungroup()

top_naive_CD4_markers %>%  group_by(cluster) %>%  top_n(n = 10, wt = avg_log2FC) -> top10_RNA_hard_naive_CD4_markers


heatmap1_c <- DoHeatmap(naive_CD4_after, features = top10_RNA_naive_CD4_markers$gene, slot="scale.data", assay="RNA", size=3) + theme(text=element_text(size=10))

# hard-clustering
heatmap2_c <- DoHeatmap(naive_CD4_after, features = top10_RNA_hard_naive_CD4_markers$gene, slot="scale.data", assay="RNA", size=3) + theme(text=element_text(size=10))
```

### Visualization of improved heatmap

```{r}
DefaultAssay(naive_CD4_after) <- "RNA"
naive_CD4_after_FOR_HEATMAP = naive_CD4_after
naive_CD4_after_FOR_HEATMAP <- ScaleData(naive_CD4_after_FOR_HEATMAP, 
                           features = top_naive_CD4_markers$gene, 
                           vars.to.regress= c("nCount_RNA"))
```

```{r}
DoHeatmap(naive_CD4_after_FOR_HEATMAP, features = top10_RNA_hard_naive_CD4_markers$gene, slot="scale.data", assay="RNA", size=3) + theme(text=element_text(size=10))
```

# Subset cluster 3 in naive_CD4_T

```{r}
naive_CD4_after_c012 <- subset(naive_CD4_after, subset = seurat_clusters != 3 )
```

```{r}
# normalize RNA count data first
naive_CD4_after_c012 <- NormalizeData(naive_CD4_after_c012)

# find variable features per patient
naive_CD4_after_c012.list <- SplitObject(naive_CD4_after_c012, split.by = "patient")
for (i in 1:length(naive_CD4_after_c012.list)) {
  naive_CD4_after_c012.list[[i]] <- FindVariableFeatures(naive_CD4_after_c012.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = TRUE)
}

# Return top scoring 2,000 features
features_naive_CD4_after_c012 <- SelectIntegrationFeatures(naive_CD4_after_c012.list)

rm(naive_CD4_after_c012.list)
```

```{r}
naive_CD4_after_c012 <- ScaleData(naive_CD4_after_c012, 
                           features = features_naive_CD4_after_c012, 
                           vars.to.regress= c("nCount_RNA", "S.Score", "G2M.Score")
                           )
```

```{r}
naive_CD4_after_c012 = RunPCA(naive_CD4_after_c012, npcs = 40, verbose = TRUE , features = features_naive_CD4_after_c012)
```

```{r}
naive_CD4_after_c012 <- RunHarmony.Seurat(naive_CD4_after_c012, "patient", 
                                          reduction.use  = "pca",
                                          reduction.save = "harmony",
                                          dims.use = 1:30)

naive_CD4_after_c012 <- naive_CD4_after_c012 %>%
    RunUMAP(reduction = "harmony", dims= 1:30, n.neighbors = 20, min.dist = 0.2)

naive_CD4_after_c012 <- naive_CD4_after_c012 %>%
  FindNeighbors(reduction = "harmony" , dims = 1:30) %>%
    FindClusters(resolution = 0.2)
```

```{r}
DimPlot(naive_CD4_after_c012, reduction = "umap", pt.size = 0.05)

DimPlot(naive_CD4_after_c012, reduction = "umap", pt.size = 0.05 , group.by = "patient")
DimPlot(naive_CD4_after_c012, reduction = "umap", pt.size = 0.05 , group.by = "timepoint")

# dittoDimPlot(C0_healthy, "DICE", reduction.use = "umap", do.letter = FALSE, do.label=F, split.by = "DICE")

FeaturePlot(naive_CD4_after_c012, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo", "S.Score","G2M.Score") , coord.fixed = T, ncol= 3, reduction = "umap")


FeaturePlot(naive_CD4_after_c012, features = c("ADGRE5", "TCF7", "BCL2", "IL7R") , coord.fixed = T, ncol= 3, reduction = "umap")

VlnPlot(naive_CD4_after_c012, c("ADGRE5", "TCF7", "BCL2", "IL7R"), group.by =  "timepoint", pt.size=0)

VlnPlot(naive_CD4_after_c012, c("ADGRE5", "TCF7", "BCL2", "IL7R"), group.by =  "seurat_clusters", pt.size=0)
```

### Cluster Proportion

```{r}
pt <- table(naive_CD4_after_c012$seurat_clusters, naive_CD4_after_c012$timepoint)
pt <- as.data.frame(pt)
pt$Var1 <- as.character(pt$Var1)

# pt$Var2 <- factor(pt$Var2, levels = c("T01", "HD"))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  #scale_fill_manual(values = brewer.pal(12, "Paired")) +
  scale_fill_manual(values = hue_pal()(length(unique(naive_CD4_after$seurat_clusters))) )   +
  theme(legend.title = element_blank())
```

### Marker selection

```{r}
naive_CD4_after_c012_RNA_marker <- FindAllMarkers(naive_CD4_after_c012, 
                                           logfc.threshold = 0.1,
                                           only.pos = TRUE,
                                           min.pct = 0.1,
                                           assay = "RNA")
```

```{r}
# heatmap
naive_CD4_after_c012_RNA_marker %>%  group_by(cluster) %>%  top_n(n = 10, wt = avg_log2FC) -> top10_RNA_naive_CD4_after_c012_markers

# gene hard clustering
top_naive_CD4_after_c012_markers <- top10_RNA_naive_CD4_after_c012_markers %>%
  group_by(gene) %>%
  top_n(n = 1, wt = avg_log2FC) %>%
  ungroup()

top_naive_CD4_after_c012_markers %>%  group_by(cluster) %>%  top_n(n = 10, wt = avg_log2FC) -> top10_RNA_hard_naive_CD4_after_c012_markers


heatmap1_c <- DoHeatmap(naive_CD4_after_c012, features = top10_RNA_naive_CD4_after_c012_markers$gene, slot="scale.data", assay="RNA", size=3) + theme(text=element_text(size=10))

# hard-clustering
heatmap2_c <- DoHeatmap(naive_CD4_after_c012, features = top10_RNA_hard_naive_CD4_after_c012_markers$gene, slot="scale.data", assay="RNA", size=3) + theme(text=element_text(size=10))
```

```{r}
DefaultAssay(naive_CD4_after_c012) <- "RNA"
naive_CD4_after_c012_FOR_HEATMAP = naive_CD4_after_c012
naive_CD4_after_c012_FOR_HEATMAP <- ScaleData(naive_CD4_after_c012_FOR_HEATMAP, 
                           features = top_naive_CD4_after_c012_markers$gene, 
                           vars.to.regress= c("nCount_RNA"))
```

```{r}
DoHeatmap(naive_CD4_after_c012_FOR_HEATMAP, features = top10_RNA_hard_naive_CD4_after_c012_markers$gene, slot="scale.data", assay="RNA", size=3) + theme(text=element_text(size=10))
```

## Save RDS files

```{r}
# save RDS files
setwd("/home/sbm/kyungmin/HYIBB/scRNA/Data/GSE210065/Result3(after_treatment)/")
saveRDS(CHOL_HD_patient_after, "CHOL_HD_patient_after.rds")
saveRDS(naive_CD4_after, "naive_CD4_after.rds")
write.csv(naive_CD4_after_RNA_marker, "naive_CD4_RNA_marker_posttreatment.csv")
write.csv(protein_markers_HD_patient_after, "protein_markers_HD_patient_posttreatment.csv")
```

## Proportion of other T cells

```{r}
naive_CD4_after <- readRDS("/home/sbm/kyungmin/HYIBB/scRNA/Data/GSE210065/Result3(after_treatment)/naive_CD4_after.rds")
Naive_ortholog_final_mouse <- readRDS("/home/sbm/kyungmin/HYIBB/scRNA/Data/Annotation_ref/Label_transfer/Naive_ortholog_final_mouse.rds")
CHOL_HD_patient_after <- readRDS("/home/sbm/kyungmin/HYIBB/scRNA/Data/GSE210065/Result3(after_treatment)/CHOL_HD_patient_after.rds")
```

```{r}
Tcell.query <- naive_CD4_after
Tcell.anchors <- FindTransferAnchors(reference = Naive_ortholog_final_mouse, 
                                     query = Tcell.query, dims = 1:30, 
                                     k.filter = 200, reference.reduction = "pca", 
                                     scale= FALSE)
```
